---
title: "General Social Survey"
author: "Karen Gomez"
date: "Year 2023"
output: 
  html_document:
    number_sections: true
    toc: true
---


```{r include=FALSE,echo=FALSE}
require(tigerstats)
require(tidyverse)
require(foreign)
source("GrabData.R")
```




#  Question 1

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data in the dataframe GSSdata is from the 2018 General Social Survey. The first blocks of R-code has selected down a subset of the data to just 16 variables. It has further removed unwanted factor levels in much of the data. Examine the code in the GrabData.R file to see what it is doing. Some of the variables are categorical and others are numerical. Be sure to do a variable analysis before tackling each question.  
First question - Is opinion on the death penalty (CAPPUN) independent of gun ownership (OWNGUN)?

$H_0$: There is no relationship between the opinion of death penalty & gun ownership. They are independent.

$H_A$: There is a relationship between the opinion of death penalty & gun ownership. They are dependent.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Both variables are categorical; each with 2 levels. Thus, we will use a chi-squared and a fisher xy test for odd ratios. 

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
barchartGC(~CAPPUN + OWNGUN,data=GSSdata)
barchartGC(~CAPPUN + OWNGUN,data=GSSdata, type="percent")

```

There is a clear difference. The pink is in favor by 60. While the blue is in favor by 55. The pink is opposed by 45. While the blue is opposed by about 20. The pink is in favor more by 5. While the blue is more opposed by 25. 

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~CAPPUN + OWNGUN, data=GSSdata)
rowPerc(table1)
colPerc(table1)
```

This table shows that if you believe in capital punishment you are likely not in favor of either yes or no. If you are opposed to capital punishment you are likely to say no 71% of the time. This table shows that if you are 18% more likely to say no

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for table1 is 0.02022. The chi square statistic is 5.3932 with a degree of freedom of 1. The chi-square statistic p-value is .01651 which is the same p-value as the fisher exact test. The odds ratio is 2.27 to 1. Therefore the odds are 2.27 to 1 likelier that somebody who has a gun will be in favor of capital punishment. We can reject the null hyoithesis because the p-value is too small.

#  Question 2

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Is there a difference in years of education (EDUC) between the those who did or did not vote in 2012 (VOTE12)?

$H_0$: There is no difference between years of education between those who did or didn't vote in 2012.

$H_A$: There is a difference between years of education between those who did or didn't vote in 2012.

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

There is a Cat vs. Num. Since the difference in years is numerical & if they did or didn't vote is categorical. We will use a box plot, favstats, and a t.test for this question.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
plot(EDUC~VOTE16,data=GSSdata)
```

The median for the voted category is 15. The median for the did not vote category is 13. There are 3 outliers in the did not vote category. The 1.5 IQR for the voted box plot is a lot lengthier than the did not vote. The IOR for the Voted category is much higher than the did not voted. Currently, there is clearly a difference between the two.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(EDUC~VOTE16,data=GSSdata)
```

The voted category has a mean of 15.625 and a median of 15. The did not vote category has a mean of 13.68 and a median is 13. This clearly states that there is a difference between the two even more. The n value for the voted category is 125. The did not vote category is 50.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
t.test(EDUC~VOTE16,data=GSSdata)
```

The t-value is 4.598. The degrees of freedom is 102.27. The p-value 15 0.00001224. Thus, we can reject the null hypothesis since the p-value is too small. There is clearly a difference between the two variables.

#  Question 3

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

Determine if a persons performance on the wordsum test (WORDSUM) is independent of their level of schooling (DEGREE).

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  

$H_A$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE 

## Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

WORDSUM is a numerical variable, DEGREE is a categorical variable.

##Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

### Descriptive Results

#### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
boxplot(WORDSUM~DEGREE,data=GSSdata)
```

From the box plot it appears that, the students in less than high school have a median less than that of the students in high school and in Bachelors.

#### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSSdata)
```

From this table it appears that the median of the Graduate students is the greatest with 11. On the other hand there is a difference in the ability in the WORDSUM.

### Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSSdata)
summary.aov(model1)
```

The P-value is 0.000236. Because the P-value is less than 0.05, we reject the null hypothesis. 

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

